---
title: "Baby-Names"
author: "Fanyi Zeng"
date: "4/23/2022"
output: github_document
---

Data source: http://hadley.github.io/babynames/

Data description:
year: double, Year of birth
sex: character, Binary sex of the baby
name: character, Name of the baby
n: integer, Raw count
prop: double, Proportion of total births for that year


```{r install data}
babynames <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-03-22/babynames.csv')
```

```{r load library, message=FALSE}
library(tidyverse)
```

Let's see the top names of 1880.

For girls, top 10 names were Mary, Anna, Emma, Elizabeth, Minnie, Margaret, Ida, Alice, Bertha, and Sarah.

For boys, top 10 names were John, William, James, Charles, George, Frank, Joseph, Thomas, Henry, and Robert.

The most enduring girls' name from 1880 to 2017 seems to be Emma. Boys' names are William and James.

```{r girl 1880}
babynames %>%
  filter(year == "1880" & sex == "F") %>%
  arrange(desc(n))
babynames %>%
  filter(year == "1880" & sex == "M") %>%
  arrange(desc(n))
```

Let's see the top names of 2017.

For girls, top 10 names were Emma, Olivia, Ava, Isabella, Sophia, Mia, Charlotte, Amelia, Evelyn, and Abigail. 

For boys, top 10 names were Liam, Noah, William, James, Logan, Benjamin, Mason, Elijah, Oliver, and Jacob.

```{r 2017}
babynames %>%
  filter(year == "2017" & sex == "F") %>%
  arrange(desc(n))
babynames %>%
  filter(year == "2017" & sex == "M") %>%
  arrange(desc(n))
```

The most popular names in 1800, Mary and Anna, largely decreased in popularity over the years. Emma being an enduring name, was both popular in 1800 and in 2017; it has experienced a gradual decline since 1800 and a revival since 2000. Olivia has also gained popularity around the same time since 2000.

```{r girl trend}
babynames %>%  
  filter(name == "Mary"|name == "Anna") %>%
  ggplot(aes(x=year, y=prop, color = name)) +
  geom_point() +
  geom_line()
babynames %>%  
  filter(name == "Emma"|name == "Olivia") %>%
  ggplot(aes(x=year, y=prop, color=name)) +
  geom_point() +
  geom_line()
```

Similarly, John and William became less favorable over the years. Liam became more prevalent around 2000, followed by Noah. 

```{r boy trend}
babynames %>%  
  filter(name == "John"|name == "William") %>%
  ggplot(aes(x=year, y=prop, color = name)) +
  geom_point() +
  geom_line()
babynames %>%  
  filter(name == "Liam"|name == "Noah") %>%
  ggplot(aes(x=year, y=prop, color=name)) +
  geom_point() +
  geom_line()
```

A general trend is that, names that very extremely common (representing a large %) in 1800 were much less dominant in 2017, perhaps because the population has grown more and more diverse in terms of national origin, race/ethnicity, religious belief, etc., which are factors that might potentially affect how people choose to name their babies.
